Exploring the Validity Range of PLL Analytical Models

### 2025-05-15 CaseStudies/Example3_2_LimitsoftheTheorySoFar/Report/article/article.qmd
Studying.

@@

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$

@@

@@

Using the value of $t_m$, the maximum normalized phase error for a frequency step, $\theta_{e_\text{m,n}}$, is given by:

@@

As a result, the maximum normalized phase error is approximately `{python} round(theta_e_m_n, 2)`.

@@

$$
\begin{split}
\theta_{e_m}&=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}\implies\\
\Delta\omega_m&=\frac{\theta_{e_m\omega_n}}{\theta_{e_\text{m,n}}}
\end{split}
$$

@@

<!-- TODO: Prepare simulations to verify these responses -->

@@

# Parâmetros iniciais
N = 5
A0 = 1
KVCO = 10e6
I = 1e-3
Kphase = I / (2 * np.pi)
```

@@

# Parâmetros do PLL
C1 = (Kphase * KVCO) / (N * omega_n**2)
C2 = C1 / 10
R1 = (zeta * 4 * np.pi * N * omega_n) / (I * KVCO)

@@

<!--
```{python}
#| echo: false

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import lsim, TransferFunction

# t_s teórico (1%)
t_s_theory = np.log(0.01 * np.sqrt(1 - zeta_meas**2)) / (-zeta_meas * omega_n_meas)

# Função de transferência
#K = A0 * Kphase * KVCO / N
#F_num = [C1 * R1, 1]
#F_den = [C1 * C2 * R1, C1 + C2, 0]
#G_num = [K * C1 * R1, K]
#G_den = [C1 * C2 * R1, C1 + C2, 0, 0]
#H_num = G_num
#H_den = np.polyadd(G_den, G_num)
#H_sys = TransferFunction(H_num, H_den)

# Simulação com entrada rampa de fase
Delta_omega = step_m_ang  # rad/s
t0 = 0
t_final = 20e-6
dt = 0.1e-6
t = np.arange(0, t_final + dt, dt)
theta_R = Delta_omega * (t - t0)
theta_R[t < t0] = 0
_, y, _ = lsim(H_sys, U=theta_R, T=t)

# Determinar valor final estimado da saída (últimos 10% dos pontos)
final_value = theta_R[-1]

# Tolerância de 1%
tolerance = 0.01 * final_value

# Cálculo da tolerância com base em 1% da entrada rampa a cada instante
within_bounds = np.abs(y - theta_R) <= tolerance

# Encontrar primeiro instante em que y entra e permanece dentro de ±1%
settling_index = None
for i in range(len(y)):
    if np.all(within_bounds[i:]):
        settling_index = i
        break

settling_time = t[settling_index] if settling_index is not None else None

# Plotagem
plt.figure(figsize=(10, 6))
plt.plot(t * 1e6, y, 'b', label='Output', linewidth=1.5)
plt.plot(t * 1e6, theta_R, 'r--', label='Input', linewidth=1)
if settling_time is not None:
    plt.axvline(settling_time * 1e6, color='g', linestyle='--', label=f'Settling Time ≈ {settling_time*1e6:.2f} µs')
plt.plot(t * 1e6, theta_R + tolerance, 'k--', linewidth=0.8, label='±1% Band')
plt.plot(t * 1e6, theta_R - tolerance, 'k--', linewidth=0.8)
plt.xlabel('Time [µs]')
plt.ylabel('Phase [rad]')
plt.title('PLL Phase Response to a Frequency Step at t₀\n(With 1% Settling Time)')
plt.legend()
plt.grid(True)
plt.tight_layout()
plt.show()

settling_time, t_s_theory

```
-->

@@

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas
u = np.ones_like(t) * step_amplitude

@@

Com o PLL de segunda-ordem devidamente definido, a função de transferência desse sistema é dada por:

@@

foi aplicado um step em fase com a mesma amplitude do step em frequência calculado anteriormente, ou seja, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang)`rad/s. Aplicando esse sinal à [Eq. @eq-transfer-function], obtém-se a resposta desse sistema ao degrau em fase.

>>>>

For the analysis of the linear transient behavior of a PLL, the phase error, rather than the output phase, is needed. So, from the [Fig. @fig-PLL-integerN-blocks], it's possible to get the relation between the phase error, $\theta_e$, and the input phase reference, $\theta_R$. This transfer function can be expressed as:

$$
\begin{split}
\frac{\theta_e}{\theta_R}&=\frac{s^2}{s^2+2\zeta\omega_ns+\omega_n^2}
\end{split}
$$ {#eq-phase-error-transfer-function}

![Block diagram of Integer-N](../images/PLL_integerN_blocks.png){#fig-PLL-integerN-blocks}

The example whose the resolution is being given here is studying the effect of an input frequency step. Since the input is described by phase, as it's possible to be seen in [Eq. @eq-phase-error-transfer-function] ($\theta_R$), it's possible to take the phase equivalent of a frequency step, which is a phase ramp, since the phase is the integral of frequency, and the integral of a step is a ramp. Thus, the input can be described as:

$$
\theta_R=\frac{\Delta\omega}{s^2}
$$ {#eq-phase-of-frequency-step}

Multiplying the reference phase of [Eq. @eq-phase-of-frequency-step] to the [Eq. @eq-phase-error-transfer-function], results in:

$$
\begin{split}
\theta_e&=\frac{\Delta\omega}{s^2+2\zeta\omega_ns+\omega_n^2}
\end{split}
$$

And getting the inverse Laplace transform, for $\zeta<1$, results in:

$$
\theta_e(t)=\Delta\omega\cdot\frac{1}{\omega_n\sqrt{1-\zeta^2}}e^{-\zeta\omega_n t}\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)
$$ {#eq-inverse-laplace-phase-error}

@@

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$ {#eq-instant-maximum-frequency-step}

@@

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which [Eq. @eq-inverse-laplace-phase-error] reaches its maximum value.

@@

First of all, instead of using the [Eq. @eq-inverse-laplace-phase-error] and apply $t_m$, it's calculated the maximum normalized phase error, which is the same equation normalized by the input signal ($\Delta\omega$) and the natural frequency ($\omega_n$). So the maximum normalized phase error, $\theta_{e_\text{m,n}}$, is given by:

@@

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

@@

$$
\begin{split}
\theta_{e_m}&=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}\implies\\
\Delta\omega_m&=\frac{\theta_{e_m}\omega_n}{\theta_{e_\text{m,n}}}
\end{split}
$$

@@

@@

# Parâmetros iniciais
N = 5
A0 = 1
KVCO = 10e6
I = 1e-3
Kphase = I / (2 * np.pi)
K = A0 * Kphase * KVCO / N
```

@@

# Parâmetros do PLL
C1 = (Kphase * KVCO) / (N * omega_n**2)
C2 = C1 / 10
R1 = (zeta * 4 * np.pi * N * omega_n) / (I * KVCO)
```

@@

@@

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = 2*step_m_ang_meas
u = np.ones_like(t) * step_amplitude

@@

With the second-order PLL properly defined, the transfer function of this system is given by:

@@

A step with the same amplitude as the previously calculated frequency step was also applied in phase, that is, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang_meas)`rad/s. Applying this signal to [Eq. @eq-transfer-function], the response of this system to the phase step is obtained. With this step, the settling time achieved is $t_s=$ `{python} format(t_s_meas)`s. Through simulation, it is possible to verify that the obtained value is $t_s=$ `{python} format(ts_graphical)` s. The simulated response can be visualized in [Fig. @fig-pll-step-response].
